SSwAB-IT Phase 1 Analysis

### 2025-09-04 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
added PABAK, etc.

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(eulerr)
library(ggpubr)
```

@@

unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,

@@

df <- full_join(df,

@@

db_demographics <- left_join(db_demographics,
  db_clinic_obtained_values %>% select(record_id, cos_collection_date),

@@

  as.numeric(difftime(db_demographics$cos_collection_date,

@@

  )) / 365.25

@@

db_demographics$age_group <- cut(db_demographics$age_at_cos,

@@

    record_id, who_collected_cos, who_collected_the_sos,

@@

  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"

@@

db_demographics$Race <- factor(db_demographics$Race,

@@

    "White", "Other", "Multiple Races", "American Indian or Alaskan Native"

@@

db_demographics$Ethnicity <- factor(db_demographics$Ethnicity,

@@

db_medications <- left_join(db_medications,

@@

  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",

@@

  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"

@@

df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")

@@

df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")

@@

df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")

@@

df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")

@@

df <- left_join(df, db_demographics %>% select(record_id, group, age_group),

@@

```{r}
db_demographics %>%
  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = group, missing_text = "Missing",
    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
    label = list(
      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
    )
  ) %>%
  add_overall() %>%
  add_p(
    test = list(c("Race", "Ethnicity") ~ "fisher.test"),
    test.args = age_group ~ list(workspace = 2e9)
  ) %>%
  separate_p_footnotes() %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt()
```

@@

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  colnames(d) <- c("COS", "SOS")
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix(d$SOS,
    reference = d$COS,
    positive = "Positive", mode = "everything",
    dnn = c("SOS", "COS")
  )
  # Print
  cat("\n")
  cat("\n")
  cat("####", "2x2 Table")
  cat("\n")
  cat("\n")
  d %>%
    tbl_cross(percent = "cell") %>%
    print()
  cat("\n")
  cat("\n")
  cat("####", "2x2 Table (With CIs)")
  cat("\n")
  cat("\n")
  d %>%
    tbl_summary(by = COS, percent = "cell") %>%
    add_ci(method = list(all_categorical() ~ "wald")) %>%
    print()
  cat("\n")
  cat("\n")
  cat("Overall accuracy (both negative/negative and positive/positive) was ",
    scales::percent(c$overall["Accuracy"], accuracy = 0.1), " (95% CI: ",
    scales::percent(c$overall["AccuracyLower"], accuracy = 0.1), " - ",
    scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1), ").",
    sep = ""
  )
  cat("\n")
  cat("\n")
  cat("Kappa was ", round(c$overall["Kappa"], 3), ".", sep = "")
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of negative results")
  cat("\n")
  cat("\n")
  print(plot(euler(d == "Negative"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
  ))
  cat("\n")
  cat("\n")
  cat("#### Euler diagram of positive results")
  cat("\n")
  cat("\n")
  print(plot(euler(d == "Positive"),
    quantities = list(cex = 3),
    fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
  ))
  cat("\n")
  cat("\n")
}
agreement_test <- function(bug, group = "group", data = df) {
  # Prepare data
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos, group)]
  d <- d[complete.cases(d), ]
  colnames(d) <- c("COS", "SOS", "Group")
  # Agreement and detailed agreement
  d$agree <- d$COS == d$SOS
  d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
  d$detailed_agree <- factor(d$detailed_agree,
    levels = c(
      "COS Negative/SOS Negative", "COS Negative/SOS Positive",
      "COS Positive/SOS Negative", "COS Positive/SOS Positive"
    ),
    labels = c("COS-/SOS-", "COS-/SOS+", "COS+/SOS-", "COS+/SOS+")
  )
  # Print
  cat("\n")
  cat("\n")
  d %>%
    select(agree, detailed_agree, Group) %>%
    tbl_summary(by = Group, missing_text = "Missing", label = list(
      agree ~ "Agree (+/+ and -/-)",
      detailed_agree ~ "Detailed Agreement"
    )) %>%
    add_overall() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e9)) %>%
    separate_p_footnotes() %>%
    modify_footnote(everything() ~ NA) %>%
    print()
  cat("\n")
  cat("\n")
}
```

@@

```{r}
# All pathogens
d <- df[, c("overall_cos", "overall_sos")]
colnames(d) <- c("COS", "SOS")
d <- d[complete.cases(d), ]
a1 <- plot(euler(d == "Positive"),
  quantities = list(cex = 3),
  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
)
a2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
# MRSA
d <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
colnames(d) <- c("COS", "SOS")
d <- d[complete.cases(d), ]
b1 <- plot(euler(d == "Positive"),
  quantities = list(cex = 3),
  fills = list(fill = c("#FFC107", "#1E88E5", "#004D40"))
)
b2 <- ggplot(as.data.frame(table(d$COS, d$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
p <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
  bgcolor("White")
ggsave("/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
  plot = p, width =9, height = 6, units = "in"
)
```

>>>>

```{r setup}
#| include: false
library(keyring)
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(epiR)
library(eulerr)
library(ggpubr)
home_dir <- switch(Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT",
    "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

token <- key_get("SSwAB-IT (New)")
rcon <- redcapConnection(
    token = token,

@@

df <- full_join(
    df,

@@

db_demographics <- left_join(
    db_demographics,
    db_clinic_obtained_values %>%
        select(record_id, cos_collection_date, ppfev1),

@@

    as.numeric(difftime(
        db_demographics$cos_collection_date,

@@

    )) /
        365.25

@@

db_demographics$age_group <- cut(
    db_demographics$age_at_cos,

@@

        record_id,
        who_collected_cos,
        who_collected_the_sos,

@@

    "American Indian or Alaskan Native",
    "Asian",
    "Hawaiian or Pacific Islander",
    "Black or African American",
    "White",
    "Unknown",
    "Other"

@@

db_demographics$Race <- factor(
    db_demographics$Race,

@@

        "White",
        "Other",
        "Multiple Races",
        "American Indian or Alaskan Native"

@@

db_demographics$Ethnicity <- factor(
    db_demographics$Ethnicity,

@@

db_medications <- left_join(
    db_medications,
db_clinic_obtained_values <- left_join(
    db_clinic_obtained_values,

@@

    "staphylococcus_aureus",
    "haemophilus_influenza",
    "pseudomonas_aeruginosa",
    "pseudomonas_non_aero",
    "s_maltophilia",
    "burkholderia",
    "aspergillus",
    "ntm",

@@

    "S. aureus",
    "H. influenza",
    "P. aeruginosa",
    "Pseudomonas non-aeruginosa",
    "S. maltophilia",
    "Burkholderia",
    "Aspergillus",
    "NTM",
    "Other"

@@

df[, paste0(bugs, "_cos")] <- lapply(
    df[, paste0(bugs, "_cos")],
    factor,
    levels = c("No", "Yes"),
    labels = c("Negative", "Positive")

@@

df[, paste0(bugs, "_sos")] <- lapply(
    df[, paste0(bugs, "_sos")],
    factor,
    levels = c("No", "Yes"),
    labels = c("Negative", "Positive")

@@

df$overall_cos <- factor(
    df$overall_cos,
    levels = c(F, T),
    labels = c("Negative", "Positive")

@@

df$overall_sos <- factor(
    df$overall_sos,
    levels = c(F, T),
    labels = c("Negative", "Positive")

@@

df <- left_join(
    df,
    db_demographics %>% select(record_id, group, age_group),

@@

```{r}
db_demographics %>%
    select(group, age_at_cos, age_group, ppfev1, sex, Race, Ethnicity) %>%
    tbl_summary(
        by = group,
        missing_text = "Missing",
        statistic = all_continuous() ~ c("{median} ({p25}, {p75})"),
        label = list(
            "sex" = "Sex",
            "age_at_cos" = "Age at COS",
            "age_group" = "Age Group"
        )
    ) %>%
    add_overall() %>%
    add_p(
        test = list(c("Race", "Ethnicity") ~ "fisher.test"),
        test.args = age_group ~ list(workspace = 2e9)
    ) %>%
    separate_p_footnotes() %>%
    as_gt()
```

## By age group

```{r}
db_demographics %>%
    mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
    select(group, age_group, ppfev1, sex, Race, Ethnicity) %>%
    tbl_summary(
        by = age_group,
        missing_text = "Missing",
        statistic = all_continuous() ~ c("{median} ({p25}, {p75})"),
        label = list(
            "sex" = "Sex",
            "age_at_cos" = "Age at COS"
        )
    ) %>%
    add_overall() %>%
    add_p(
        test = list(c("Race", "Ethnicity") ~ "fisher.test")
    ) %>%
    separate_p_footnotes() %>%
    as_gt()

db_demographics %>%
    mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
    select(age_group, ppfev1) %>%
    drop_na() %>%
    ggplot(., aes(x = ppfev1)) +
    geom_histogram() +
    facet_wrap(~age_group)
```

@@

```{r}
agreement <- function(bug, data = df) {
    cos <- paste0(bug, "_cos")
    sos <- paste0(bug, "_sos")
    d <- data[, c(cos, sos)]
    colnames(d) <- c("COS", "SOS")
    d <- d[complete.cases(d), ]
    # Agreement metrics
    c <- confusionMatrix(
        d$SOS,
        reference = d$COS,
        positive = "Positive",
        mode = "everything",
        dnn = c("SOS", "COS")
    )
    pabak <- epi.kappa(table(d$COS, d$SOS))
    # Print
    cat("\n")
    cat("\n")
    cat("####", "2x2 Table")
    cat("\n")
    cat("\n")
    d %>%
        tbl_cross(percent = "cell") %>%
        print()
    cat("\n")
    cat("\n")
    cat("####", "2x2 Table (With CIs)")
    cat("\n")
    cat("\n")
    d %>%
        tbl_summary(by = COS, percent = "cell") %>%
        add_ci(method = list(all_categorical() ~ "wald")) %>%
        print()
    cat("\n")
    cat("\n")
    cat(
        "Overall accuracy (both negative/negative and positive/positive) was ",
        scales::percent(c$overall["Accuracy"], accuracy = 0.1),
        " (95% CI: ",
        scales::percent(c$overall["AccuracyLower"], accuracy = 0.1),
        " - ",
        scales::percent(c$overall["AccuracyUpper"], accuracy = 0.1),
        ").",
        sep = ""
    )
    cat("\n")
    cat("\n")
    cat(
        "Kappa was ",
        round(pabak$kappa$est, 3),
        " (95% CI: ",
        round(pabak$kappa$lower, 3),
        " - ",
        round(pabak$kappa$upper, 3),
        ")",
        ".",
        sep = ""
    )
    cat("\n")
    cat("\n")
    cat(
        "Prevalence and Bias Adjusted Kappa (PABAK) was ",
        round(pabak$pabak$est, 3),
        " (95% CI: ",
        round(pabak$pabak$lower, 3),
        " - ",
        round(pabak$pabak$upper, 3),
        ")",
        ".",
        sep = ""
    )
    cat("\n")
    cat("\n")
    cat("#### Euler diagram of negative results")
    cat("\n")
    cat("\n")
    print(plot(euler(d == "Negative"), quantities = list(cex = 3)))
    cat("\n")
    cat("\n")
    cat("#### Euler diagram of positive results")
    cat("\n")
    cat("\n")
    print(plot(euler(d == "Positive"), quantities = list(cex = 3)))
    cat("\n")
    cat("\n")
}
agreement_test <- function(bug, group = "group", data = df) {
    # Prepare data
    cos <- paste0(bug, "_cos")
    sos <- paste0(bug, "_sos")
    d <- data[, c(cos, sos, group)]
    d <- d[complete.cases(d), ]
    colnames(d) <- c("COS", "SOS", "Group")
    # Agreement and detailed agreement
    d$agree <- d$COS == d$SOS
    d$detailed_agree <- paste0("COS ", d$COS, "/", "SOS ", d$SOS)
    d$detailed_agree <- factor(
        d$detailed_agree,
        levels = c(
            "COS Negative/SOS Negative",
            "COS Negative/SOS Positive",
            "COS Positive/SOS Negative",
            "COS Positive/SOS Positive"
        ),
        labels = c("COS-/SOS-", "COS-/SOS+", "COS+/SOS-", "COS+/SOS+")
    )
    # Print
    cat("\n")
    cat("\n")
    d %>%
        select(agree, detailed_agree, Group) %>%
        tbl_summary(
            by = Group,
            missing_text = "Missing",
            label = list(
                agree ~ "Agree (+/+ and -/-)",
                detailed_agree ~ "Detailed Agreement"
            )
        ) %>%
        add_overall() %>%
        add_p(test.args = all_tests("fisher.test") ~ list(workspace = 2e9)) %>%
        separate_p_footnotes() %>%
        modify_footnote(everything() ~ NA) %>%
        print()
    cat("\n")
    cat("\n")
}
```

@@

```{r}
# All pathogens
d_overall <- df[, c("overall_cos", "overall_sos")]
colnames(d_overall) <- c("COS", "SOS")
d_overall <- d_overall[complete.cases(d_overall), ]
a1 <- plot(euler(d_overall == "Positive"), quantities = list(cex = 3))
a2 <-
    ggplot(
        as.data.frame(table(d_overall$COS, d_overall$SOS)),
        aes(Var1, Var2)
    ) +
    geom_tile(color = "white", fill = "#1E88E5") +
    geom_text(
        aes(
            label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")
        ),
        color = "white"
    ) +
    xlab("COS") +
    ylab("SOS") +
    theme_bw() +
    theme(text = element_text(size = 15))
# MRSA
d_MSSA <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
colnames(d_MSSA) <- c("COS", "SOS")
d_MSSA <- d_MSSA[complete.cases(d_MSSA), ]
b1 <- plot(euler(d_MSSA == "Positive"), quantities = list(cex = 3))
b2 <-
    ggplot(as.data.frame(table(d_MSSA$COS, d_MSSA$SOS)), aes(Var1, Var2)) +
    geom_tile(color = "white", fill = "#1E88E5") +
    geom_text(
        aes(
            label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")
        ),
        color = "white"
    ) +
    xlab("COS") +
    ylab("SOS") +
    theme_bw() +
    theme(text = element_text(size = 15))
p1 <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
    bgcolor("White")
ggsave(
    "./Dissemination/Figures/fig.png",
    plot = p1,
    width = 9,
    height = 6,
    units = "in"
)
a1 <- plot(
    euler(d_overall == "Positive"),
    quantities = list(cex = 3),
    main = "Any CF Pathogen"
)
b1 <- plot(
    euler(d_MSSA == "Positive"),
    quantities = list(cex = 3),
    main = "MSSA"
)
p2 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
    bgcolor("White")
ggsave(
    "./Dissemination/Figures/euler_bw.png",
    plot = p2,
    width = 9,
    height = 6,
    units = "in"
)
a1 <- plot(
    venn(d_overall == "Positive"),
    quantities = list(cex = 3),
    main = "Any CF Pathogen"
)
b1 <- plot(
    venn(d_MSSA == "Positive"),
    quantities = list(cex = 3),
    main = "MSSA"
)
p3 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
    bgcolor("White")
ggsave(
    "./Dissemination/Figures/venn_bw.png",
    plot = p3,
    width = 9,
    height = 6,
    units = "in"
)
```
